---
title: "Tooth Growth Analysis"
author: "Paul Askew"
date: "21 December 2015"
output: word_document
---
##Data Processing and Exploratory Analysis##

This project analyses the toothgrowth dataset supplied with R.  Performing `?Toothgrowth` gives this description:

*The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).*

First of all we load the data and construct a simple plot to see if there is any demonstrable difference.  It is clear from the plot that there appears to be a difference so it is worthwhile analysing the data further.
```{r,warning=FALSE, echo = FALSE}
library(datasets)
data("ToothGrowth")
library(ggplot2)
#ggplot(aes(x=supp, y=len), data=ToothGrowth) + #geom_boxplot(aes(fill=supp))+ 
#    xlab("Supplement Type") +ylab("Tooth Length (mm)") 

ggplot(aes(x = supp, y = len), data = ToothGrowth) +
    geom_boxplot(aes(fill = supp)) + facet_wrap(~ dose)+ 
    xlab("Supplement Type") +ylab("Tooth Length (mm)") 
summary(ToothGrowth)
```


#Analysis

Looking at the figures, it would appear that there is a difference in toothgrowth depending on the type of supplement given. First of all let's split the data into unique sets depending upon the supplement and the dose.

```{r}
AA0_5 <- ToothGrowth[1:10, ] # ascorbic acid, 0.5 mg dose
AA1 <- ToothGrowth[11:20, ] # ascorbic acid, 1 mg dose
AA2 <- ToothGrowth[21:30, ] # ascorbic acid, 2 mg dose
OJ0_5 <- ToothGrowth[31:40, ] # orange juice, 0.5 mg dose
OJ1 <- ToothGrowth[41:50, ] # orange juice, 1 mg dose
OJ2 <- ToothGrowth[51:60, ] # orange juice, 2 mg dose
```

Now let's test the null hypothesis, ie there is no difference in effectiveness between each supplement at each dose level.  To do this we will use the t test to compare the average tooth length for each supplement at a given dose.

```{r}
t.test(OJ0_5$len, AA0_5$len, var.equal = TRUE)$conf # 0.5 mg dose
t.test(OJ1$len, AA1$len, var.equal = TRUE)$conf # 1 mg dose
t.test(OJ2$len, AA2$len, var.equal = TRUE)$conf # 2 mg dose
```

#Conclusions#

From this we can conclude that there is a difference between the two supplements, Orange Juice and Ascorbic Acid, for doses of 0.5mg and 1 mg but there is no discernible difference between doses of 2mg (as the 95% confidence interval passes through 0).  Therefore if longer toothgrowth is required, Orange Juice is the preferred supplement.

This assumes that the populations tested are independent of each other and that the distribution of results is approximately normal.
